---
title: "Group Project"
author: "Ruixin Lou, Chloe Liu, Paul McKee, Kristi Van Meter"
date: "`r Sys.Date()`"
header-includes:
   - \usepackage{dcolumn}
output:
  pdf_document:
    toc: yes
  html_document:
    self_contained: yes
    code_folding: hide
    code_download: yes
    toc: yes
    toc_float: yes
    number_sections: yes
---

```{r global options, include = FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE,include = FALSE)
```

```{r}
library(haven)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(janitor)
library(kableExtra)
library(gridExtra)
library(equatiomatic)
library(ggfortify)
library(stargazer)
library(caret)
library(table1)
library(ggcorrplot)
```

```{r setup, include=FALSE}
data = read_xpt('LLCP2021.XPT ')
```


\newpage

# Abstract

# An Analysis of CDC Data to Predict Mental Health Problems and Diabetes

Since 1984, the Behavioral Risk Factor Surveillance System (BRFSS) has collected health-related data via phone surveys across all fifty states to get a better understanding of the population’s health-related risk behaviors, chronic health conditions, and use of preventive services. The open source dataset is sent to the CDC where it is aggregated by state and published at the end of the year (CDC Behavioral Risk Factor Surveillance System Survey, 2001). Data collected includes information about how and when the interview was conducted, demographic information, medical history, current medical condition, behaviors that increase/ lower health risks, and utilization of the healthcare system. Medical conditions and history include mental and physical health conditions, as well as acute and chronic illness. Of note, given the number of different administrating centers and values/ rules within different states, not all questions are asked in every state, leading to many missing values in the CDC’s full dataset.

This project proceeded with the following two research questions: 
- R1: How well do exercise habits and medical history predict the number of poor mental health days during a 30 day period?, 
- R2: What are the relationships between different demographics (e.g., race, age, gender, healthcare status, etc.) and the likelihood of being diagnosed with diabetes?

**Subjects and Data Origin**

This report uses data provided by the Behavioral Risk Factor Surveillance System (BRFSS; CDC Behavioral Risk Factor Surveillance System Survey, 2001). The BRFSS data examined here is from the 2021 annual survey conducted by the Centers for Disease Control and Prevention (CDC). The goal of the BRFSS is to collect state-specific data on chronic diseases and conditions, access to health care, health risk behaviors, and the use of preventive health services associated with the leading causes of death and disability within the US. As such, information available includes a diverse set of factors, such as chronic health conditions, tobacco use, fruit and vegetable consumption, health care access, hypertension and cholesterol awareness, exercise habits, health status and number of healthy data, and arthritis status. The data outlined in the following report comes from 49 states, Washington D.C., Guam, Puerto Rico, and the US Virgin Islands. After being unable to collect BRFSS data over enough months to satisfy the requirements for inclusion in this 2021 dataset, Florida was excluded. All data was self-reported through interviews on the telephone. This dataset contains 303 features for a sample of 438,693 people. 


### Research Question 1: How do exercise habits and medical history predict the number of days mental health was not good in the past 30 days?

#### Methods

The first step taken to answer this first research question was to select only a subset of the features that were relevant as suggested by previous research and theory. The features selected were as follows.

Predictors: Age, Gender,Number of Personal Health Care Provider, Exercise, Arthritis, Diabetes, Asthma, Blood Pressure, Number of Days Physical Health not Good, High Cholesterol, Kidney
Outcome: Number of Days in Past 30 Days Mental Health not Good


```{r}
q1 = data%>%dplyr::select('_AGE80',BIRTHSEX,PERSDOC3,EXERANY2,HAVARTH5,DIABETE4,ASTHMA3,BPHIGH6,PHYSHLTH,TOLDHI3,CHCKDNY2,MENTHLTH)
```


```{r include = TRUE}
old_names = colnames(q1)
new_names = c("age","gender","persdoc","exercise","arthritis","diabetes","asthma","bp_high",'physical','cholesterol','kidney',"mental")
# overwrite the old names with the new names:
colnames(q1) = new_names
questions = c("Imputed Age value collapsed above 80","What was your sex at birth? Was it male or female?","Do you have one person (or a group of doctors) that you think of as your personal health care provider?","During the past month, other than your regular job, did you participate in any physical activities or exercises
such as running, calisthenics, golf, gardening, or walking for exercise?","Has a doctor, nurse or other health professional ever told you that you had some form of arthritis, rheumatoid arthritis, gout, lupus, or fibromyalgia? (Arthritis diagnoses include: rheumatism, polymyalgia rheumatic, osteoarthritis (not osteoporosis), tendonitis, bursitis, bunion, tennis elbow, carpal tunnel syndrome, tarsal tunnel syndrome, joint infection, Reiter’s syndrome, ankylosing spondylitis;)","(Ever told) (you had) diabetes? (If ´Yes´ and respondent is female, ask ´Was this only when you were pregnant?´. If Respondent says pre-diabetes or borderline diabetes, use response code 4.) ","(Ever told) (you had) asthma?","Ever Told Blood Pressure High?",'Number of Days Physical Health Not Good','Ever Told Cholesterol Is High?','Ever Told Have Kidney Disease?',"Now thinking about your mental health, which includes stress, depression, and problems with emotions, for how many days during the past 30 days was your mental health not good?")
name_combo = bind_cols(New = new_names, Old = old_names,Questions = questions)
name_combo %>% gt::gt()%>% gt::gtsave("tab_1.png", expand = 10)
```

##### Data Cleaning

For physical and mental health features, all reported data that was larger than 30 was removed. For high blood pressure and diabetes features, "yes, but only during pregnancy" and “no, pre-diabetes/high blood pressure or borderline diabetes/high blood pressure” were collapsed in binary "yes" and "no" respectively. For arthritis, persdoc, gender, asthma, bp_high, cholesterol, and kidney features "Don’t know/Not sure" responses were treated as missing data. Any data entry with missing values in any of the relevant features were excluded from analyses. This cleaning phase brought the sample down to N = 79520. With this sample size, we are sufficiently powered to answer this question so it is not a concern.


```{r}
q1_clean = q1%>%
  mutate(
    mental  = case_when(
      mental == 88 ~ 0,
      TRUE ~ mental
    ),
    physical = case_when(
      physical == 88 ~ 0,
      TRUE ~ physical
    ),
    diabetes = case_when(
      diabetes == 2 ~ 1,
      diabetes == 4 ~ 3,
      TRUE ~ diabetes
    ),
    bp_high  = case_when(
      bp_high == 2 ~ 1,
      bp_high == 4 ~ 3,
      TRUE ~ bp_high
    )
  )%>%
  filter(mental <= 30 & exercise<=2 & physical <= 30 & cholesterol<=2 & kidney <= 2 & gender <=2 & arthritis<=2 & diabetes<=3 & asthma <=2 & bp_high<=3 & persdoc<=3)
```



```{r}
q1_clean$gender <- factor(q1_clean$gender,
                        levels=c(1,2), #how it is coded in the current dataset
                        labels = c('Male','Female') #give it new names
)

q1_clean$persdoc <- factor(q1_clean$persdoc,
                        levels=c(1,2,3), #how it is coded in the current dataset
                        labels = c('only one','More than one','No') #give it new names
) 

q1_clean$exercise <- factor(q1_clean$exercise,
                        levels=c(1,2), #how it is coded in the current dataset
                        labels = c('Yes','No') #give it new names
)

q1_clean$arthritis <- factor(q1_clean$arthritis,
                        levels=c(1,2), #how it is coded in the current dataset
                        labels = c('Yes','No') #give it new names
) 

q1_clean$diabetes <- factor(q1_clean$diabetes,
                        levels=c(1,3), #how it is coded in the current dataset
                        labels = c('Yes','No') #give it new names
) 

q1_clean$asthma <- factor(q1_clean$asthma,
                        levels=c(1,2), #how it is coded in the current dataset
                        labels = c('Yes','No') #give it new names
) 

q1_clean$bp_high <- factor(q1_clean$bp_high,
                        levels=c(1,3), #how it is coded in the current dataset
                        labels = c('Yes','No') #give it new names
) 

q1_clean$cholesterol <- factor(q1_clean$cholesterol,
                        levels=c(1,2), #how it is coded in the current dataset
                        labels = c('Yes','No') #give it new names
) 

q1_clean$kidney <- factor(q1_clean$kidney,
                        levels=c(1,2), #how it is coded in the current dataset
                        labels = c('Yes','No') #give it new names
) 
```

```{r}
q1_clean = q1_clean%>%
  drop_na()
```

##### Modeling

A Poisson Regression Model (PRM) was used to regress our outcome variable of interest on the predictor variables listed above. PRM is the appropriate generalized linear model (GLM) to model count data. PRM models this data by assuming 1) the outcome variable has a Poisson distribution and 2) the logarithm of the outcome’s expected value can be adequately modeled by a linear combination of unknown parameters. Thus, as our outcome counts the number of poor mental health days in a given 30 day period, it is best modeled under this approach and assumptions. All of the aforementioned predictor variables were entered into the model at the same time to predict the logarithm of mental. Meanwhile, we also have included all the interaction terms and used backward selection using AIC as criteria to select significant interaction terms. No other transformations were made.

#### Results

##### Exploratory Data Analysis (EDA)

Descriptively, most people report experiencing no greater than 10 bad mental or physical health days out of the last 30 days. In line with previous psychological findings, most people report bad day totals that are divisible by 5. This is even more apparent when looking at the people that report greater than 10 bad days (Figure 1 left and right panel). Interestingly, there is a meaningful subsample of people who report having a total of 30 out of 30 days of poor mental or physical health. Age is relatively uniformly distributed with a slight negative skew (Figure 1 middle panel).

```{r echo = FALSE,fig.height= 2,include=TRUE,fig.width= 10, fig.cap="Histogram of Mental Health, Age and Physical Health"}
disease_toplot = q1_clean%>%
  dplyr::select(mental,age,physical)

par(mfrow = c(1, 3))
for (i in 1:ncol(disease_toplot)){
  data = as.numeric(unlist(disease_toplot[, i]))
  column_name = colnames(disease_toplot)[i]
  hist(data, col = "Steelblue",
       xlab = paste(column_name), ylab = "Count",
       main = paste(column_name," histogram"),breaks = 30)
}
```

```{r fig.height= 2,fig.width=10,include=TRUE,fig.cap="Box plots of Mental Health v.s. Gneder, Exercise and Number of Personal Health Provider"}
# Boxplot (Figure Xc)
p1 = ggplot(q1_clean, aes(x=gender, y=mental, fill=gender)) +
  geom_boxplot()+
    theme_bw() + 
  labs (x="Gender", y="Mental Health")

#Density plot (Figure Xd)
p2 = ggplot(q1_clean,aes(x=mental, fill=gender)) +
  geom_density(bw=5, alpha=0.5) +
  labs(x="Mental Health", y="Density") +
  theme_bw()

p3 = ggplot(q1_clean, aes(x=exercise, y=mental, fill=exercise)) +
  geom_boxplot()+
    theme_bw() + 
  labs (x="Exercise", y="Mental Health")

p4 = ggplot(q1_clean,aes(x=mental, fill=exercise)) +
  geom_density(bw=5, alpha=0.5) +
  labs(x="Mental Health", y="Density") +
  theme_bw()

p5 = ggplot(q1_clean, aes(x=persdoc, y=mental, fill=persdoc)) +
  geom_boxplot()+
    theme_bw() + 
  labs (x=" Personal Health Care Provider", y="Mental Health")+
  scale_x_discrete(labels=c("only one" = "1", "More than one" = ">1","No" = "0"))

p6 = ggplot(q1_clean,aes(x=mental, fill=persdoc)) +
  geom_density(bw=5, alpha=0.5) +
  labs(x="Mental Health", y="Density") +
  theme_bw()

# grid.arrange(p1, p2,p3,p4,p5,p6, ncol=6)

p2 = ggplot(q1_clean,aes(x=mental, fill=gender)) +
  geom_density(bw=5, alpha=0.5) +
  labs(x="Mental Health", y="Density") +
  theme_bw()

p4 = ggplot(q1_clean,aes(x=mental, fill=exercise)) +
  geom_density(bw=5, alpha=0.5) +
  labs(x="Mental Health", y="Density") +
  theme_bw()

p6 = ggplot(q1_clean,aes(x=mental, fill=persdoc)) +
  geom_density(bw=5, alpha=0.5) +
  labs(x="Mental Health", y="Density") +
  theme_bw()
grid.arrange(p1,p3,p5,ncol=3)
grid.arrange(p2,p4,p6, ncol=3)
```


Descriptively, there is a difference in the number of poor mental health days between male and females, with female has a larger range of unhappy days, even though the median is 0 for both (Figure 2 left panel). Of note, upon visual inspection, it appears that those that have not exercised in the past month report a greater number of poor mental health days (Figure 2 middle panel). From the right panel of Figure 2 we can observe that those with one personal health care provider report less poor mental health days than those with multiple personal health care providers or none at all, in that order. 


```{r fig.height= 2,fig.width=10,include=TRUE,fig.cap="Box plots of Mental Health vs. Arthritis, Diabetes and Asthma"}

p1 = ggplot(q1_clean, aes(x=arthritis, y=mental, fill=arthritis)) +
  geom_boxplot()+
    theme_bw() + 
  labs (x="arthritis", y="Mental Health")

p3 = ggplot(q1_clean, aes(x=diabetes, y=mental, fill=diabetes)) +
  geom_boxplot()+
    theme_bw() + 
  labs (x="diabetes", y="Mental Health")

p5 = ggplot(q1_clean, aes(x=asthma, y=mental, fill=asthma)) +
  geom_boxplot()+
    theme_bw() + 
  labs (x="asthma", y="Mental Health")

grid.arrange(p1,p3,p5,ncol=3)
```
```{r fig.height= 2,fig.width=10,include=TRUE,fig.cap="Box plots of Mental Health v.s. Gneder, Exercise and Number of Personal Health Provider"}
# Boxplot (Figure Xc)

#Density plot (Figure Xd)
p2 = ggplot(q1_clean,aes(x=mental, fill=arthritis)) +
  geom_density(bw=5, alpha=0.5) +
  labs(x="Mental Health", y="Density") +
  theme_bw()

p4 = ggplot(q1_clean,aes(x=mental, fill=diabetes)) +
  geom_density(bw=5, alpha=0.5) +
  labs(x="Mental Health", y="Density") +
  theme_bw()

p6 = ggplot(q1_clean,aes(x=mental, fill=asthma)) +
  geom_density(bw=5, alpha=0.5) +
  labs(x="Mental Health", y="Density") +
  theme_bw()

grid.arrange(p2,p4,p6, ncol=3)
# grid.arrange(p1,p3,p5,ncol=3)
```

Not surprisingly, those who suffered from arthritis (Figure 3 left panel), diabetes (middle panel), or asthma (right panel) reported greater numbers of poor mental health days compared to those who did not experience those diseases.


##### Model Results

```{r}
model = glm(mental ~ .,data = q1_clean,family = poisson)
```

```{r}
model2 = glm(mental ~ (.)^2,data = q1_clean,family = poisson)
```

```{r echo = FALSE,results='hide'}
# M1 = step(model2, scope=list(upper=model2,lower=~ age + gender + persdoc + exercise + arthritis + diabetes + asthma + bp_high + physical + cholesterol + kidney),
#          direction = "backward", 
#          trace = FALSE)
```


```{r echo = FALSE}
# drop1(M1,test='F')
```

```{r}
model3 = glm(mental ~ age + gender + persdoc + exercise + arthritis + diabetes + 
               asthma + bp_high + physical + cholesterol + kidney + age*persdoc + 
               age*exercise + age*arthritis + age*asthma + age*bp_high + age*physical+
               gender*persdoc + gender*bp_high + gender * physical + persdoc * exercise+
               persdoc * asthma + exercise*diabetes + exercise * physical+
               exercise*cholesterol+arthritis*asthma+arthritis*physical+
               diabetes*bp_high+diabetes*physical+diabetes*cholesterol+physical*kidney+
               physical*cholesterol,data = q1_clean,family = poisson)
```



```{r}
extract_eq(model3,wrap = TRUE,use_coefs = TRUE,coef_digits = 4,terms_per_line = 2)
```

$$
\begin{aligned}
\log\left[ \frac { P( \operatorname{mental} = \operatorname{1} ) }{ 1 - P( \operatorname{mental} = \operatorname{1} ) } \right] &= 3.598 - 0.0384(\operatorname{age})\ + \\
&\quad 0.427(\operatorname{gender}_{\operatorname{Female}}) + 0.1162(\operatorname{persdoc}_{\operatorname{More\ than\ one}})\ - \\
&\quad 0.2094(\operatorname{persdoc}_{\operatorname{No}}) + 0.153(\operatorname{exercise}_{\operatorname{No}})\ - \\
&\quad 0.3216(\operatorname{arthritis}_{\operatorname{No}}) - 0.0175(\operatorname{diabetes}_{\operatorname{No}})\ - \\
&\quad 0.098(\operatorname{asthma}_{\operatorname{No}}) - 0.2974(\operatorname{bp\_high}_{\operatorname{No}})\ + \\
&\quad 0.0121(\operatorname{physical}) - 0.2725(\operatorname{cholesterol}_{\operatorname{No}})\ - \\
&\quad 0.2366(\operatorname{kidney}_{\operatorname{No}}) - 0.001(\operatorname{age} \times \operatorname{persdoc}_{\operatorname{More\ than\ one}})\ + \\
&\quad 0.0103(\operatorname{age} \times \operatorname{persdoc}_{\operatorname{No}}) + 0.0033(\operatorname{age} \times \operatorname{exercise}_{\operatorname{No}})\ + \\
&\quad 0.0018(\operatorname{age} \times \operatorname{arthritis}_{\operatorname{No}}) - 0.002(\operatorname{age} \times \operatorname{asthma}_{\operatorname{No}})\ + \\
&\quad 0.0046(\operatorname{age} \times \operatorname{bp\_high}_{\operatorname{No}}) + 4e-04(\operatorname{age} \times \operatorname{physical})\ - \\
&\quad 0.0593(\operatorname{gender}_{\operatorname{Female}} \times \operatorname{persdoc}_{\operatorname{More\ than\ one}}) + 0.0119(\operatorname{gender}_{\operatorname{Female}} \times \operatorname{persdoc}_{\operatorname{No}})\ + \\
&\quad 0.0652(\operatorname{gender}_{\operatorname{Female}} \times \operatorname{bp\_high}_{\operatorname{No}}) - 0.0099(\operatorname{gender}_{\operatorname{Female}} \times \operatorname{physical})\ - \\
&\quad 0.033(\operatorname{persdoc}_{\operatorname{More\ than\ one}} \times \operatorname{exercise}_{\operatorname{No}}) - 0.2263(\operatorname{persdoc}_{\operatorname{No}} \times \operatorname{exercise}_{\operatorname{No}})\ + \\
&\quad 0.0898(\operatorname{persdoc}_{\operatorname{More\ than\ one}} \times \operatorname{asthma}_{\operatorname{No}}) - 0.0784(\operatorname{persdoc}_{\operatorname{No}} \times \operatorname{asthma}_{\operatorname{No}})\ - \\
&\quad 0.0758(\operatorname{exercise}_{\operatorname{No}} \times \operatorname{diabetes}_{\operatorname{No}}) - 0.0049(\operatorname{exercise}_{\operatorname{No}} \times \operatorname{physical})\ + \\
&\quad 0.0595(\operatorname{exercise}_{\operatorname{No}} \times \operatorname{cholesterol}_{\operatorname{No}}) - 0.14(\operatorname{arthritis}_{\operatorname{No}} \times \operatorname{asthma}_{\operatorname{No}})\ + \\
&\quad 0.0084(\operatorname{arthritis}_{\operatorname{No}} \times \operatorname{physical}) - 0.1395(\operatorname{diabetes}_{\operatorname{No}} \times \operatorname{bp\_high}_{\operatorname{No}})\ + \\
&\quad 0.0033(\operatorname{diabetes}_{\operatorname{No}} \times \operatorname{physical}) + 0.0762(\operatorname{diabetes}_{\operatorname{No}} \times \operatorname{cholesterol}_{\operatorname{No}})\ + \\
&\quad 0.0111(\operatorname{physical} \times \operatorname{kidney}_{\operatorname{No}}) + 0.0032(\operatorname{physical} \times \operatorname{cholesterol}_{\operatorname{No}})\ + \\
&\quad \epsilon
\end{aligned}
$$

Our model indicated relationships for all our predictor variables on the log counts of poor mental health days except for gender x persdoc (no vs one). We could not reject the null hypothesis for gender x persdoc (no vs one) (b = 0.01, p > 0.05, 95% CI [-0.01, 0.04]). See Table 1 below for full results.

The predictors with the largest relationships with number of poor mental health days include gender (b = 0.43, p < 0.01, 95% CI [0.41, 0.44]), arthritis (b = -0.32, p < 0.01, 95% CI [-0.35, -0.29]), high blood pressure (b = -0.30, p < 0.01, 95% CI [-0.33, -0.26]), high cholesterol (b = -0.27, p < 0.01, 95% CI [-0.29, -0.25]), kidney (b = -0.24, p < 0.01, 95% CI [-0.26, -0.21]), persdoc x exercise (b = -0.23, p < 0.01, 95% CI [-0.25, -0.20]), exercise (b = 0.15, p < 0.01, 95% CI [0.11, 0.19]), arthritis x asthma (b = -0.14, p < 0.01, 95% CI [-0.16, -0.12]), diabetes x high blood pressure (b = -0.14, p < 0.01, 95% CI [-0.16, -0.12]), persdoc (more than one vs one) (b = 0.12, p < 0.01, 95% CI [0.08, 0.15]) and asthma (b = -0.10, p < 0.01, 95% CI [-0.13, -0.06]). 

Being a female is associated with having 1.34 times more poor mental health days compared to being a male. Likewise, not exercising is associated with having 1.26 times more poor mental health days compared to those that exercise. In a similar direction, having no personal health care provider is associated with 1.15 times more poor mental health days compared to people that have exactly one. Not having arthritis or diabetes are protective factors that are associated with a reduction of 13% and 10% in the number of poor mental health days, respectively, compared to people with arthritis or diabetes.


##### Model Evaluation


**Independence** 

Data collection is conducted separately by each state. The design uses state-level, random digit dialed probability samples of the adult (ages 18 and older) population. All projects use a disproportionate stratified sample design except for Guam, Puerto Rico, and the U.S. Virgin Islands who use a simple random sample design. Since the sampling procedure conform with random sampling, we can conclude that our each of the observations in our dataset is independent.

**Multicollinearity**

```{r include = FALSE}
library(car)
vif = vif(model3)
```

Computing the VIF for all predictors, we find that all the predictors are above the threshold that indicate no multicollinearity (e.g., 1). However, since most of $GVIF^(1/(2*Df))$ values remain below 5, only exercise has a VIF slightly greater than 5, there is not a major concern that multicollinearity will introduce any meaningful issues in our data analysis and inference. 

**Residuals and Predictions**


```{r include=TRUE,fig.height=2}
#residuals and predictions
resid <- resid(model3, type = "pearson")
pred <- predict(model3,type="response")

#residuals vs fitted
qplot(y=resid, x=pred, geom="point",
      xlab = "Predicted Counts", ylab = "Pearson Residuals")
```


When plotting the residuals against the predicted values, there is a slight pattern, with lower predicted values having larger residuals and higher predicted values having smaller residuals. Since it is relatively minor, it is not a concern for the analysis or interpretation of these results.

**Overall Model Fit**

Comparing with null model:

- Model 1: mental ~ 1
- Model 2: mental ~ age + gender + persdoc + exercise + arthritis + diabetes + 
    asthma + bp_high + physical + cholesterol + kidney + age * 
    persdoc + age * exercise + age * arthritis + age * asthma + 
    age * bp_high + age * physical + gender * persdoc + gender * 
    bp_high + gender * physical + persdoc * exercise + persdoc * 
    asthma + exercise * diabetes + exercise * physical + exercise * 
    cholesterol + arthritis * asthma + arthritis * physical + 
    diabetes * bp_high + diabetes * physical + diabetes * cholesterol + 
    physical * kidney + physical * cholesterol

```{r}
nullmod = glm(mental~1,data = q1_clean,family = "poisson")
anova(nullmod,model3,test = "Chisq")
```

```{r}
anova(nullmod,model3,test = "Chisq") %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped"))
```



Comparing with model without any interaction term:

- Model 1: mental ~ age + gender + persdoc + exercise + arthritis + diabetes + 
    asthma + bp_high + physical + cholesterol + kidney
- Model 2: mental ~ age + gender + persdoc + exercise + arthritis + diabetes + 
    asthma + bp_high + physical + cholesterol + kidney + age * 
    persdoc + age * exercise + age * arthritis + age * asthma + 
    age * bp_high + age * physical + gender * persdoc + gender * 
    bp_high + gender * physical + persdoc * exercise + persdoc * 
    asthma + exercise * diabetes + exercise * physical + exercise * 
    cholesterol + arthritis * asthma + arthritis * physical + 
    diabetes * bp_high + diabetes * physical + diabetes * cholesterol + 
    physical * kidney + physical * cholesterol

```{r}
anova(model,model3,test = "Chisq") %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped"))
```

Comparing with both null model and model without any interaction term, both deviance tests showed a rather small p value (<2.2e-16) which means the tests are significant at a 0.05 significance level. We can conclude that our final model performs better than both null model and model without interaction terms. Our model overall is a good fit.

**Outliers**

```{r echo=FALSE,fig.height=2,fig.width=10}
q1_clean$cooksD <- cooks.distance(model3)
plot(q1_clean$cooksD)
```

There is no influential points or outliers in our model.

# Conclusion and Limitation

To answer the first research question, "How well do exercise habits and medical history predict the number of poor mental health days during a 30-day period?", we performed Poisson regression on the number of poor mental health days during a 30-day period.
Based on the model result, exercise indeed has positive effects on mental health, and not having arthritis or diabetes will contribute to decreasing the number of days that mental health is not good. Taylor et al.’s study (1985) provided evidence suggesting that physical activity and exercise can help reduce anxiety symptoms, enhance self-esteem, social skills, and cognitive functioning, and modify coronary-prone behavior and the body's physiological reaction to stress. In terms of diseases, many studies have shown clearly that people with arthritis with the highest pain levels are the most likely to be anxious or depressed, and people with diabetes are 2 to 3 times more likely to have depression than people without diabetes (Centers for Disease Control and Prevention, 2021, Centers for Disease Control and Prevention, 2022). Even though our model shows whether or not having kidney disease or high cholesterol is insignificant, we still cannot conclude that those diseases have no effects on mental health. Since we have deleted a large amount of data entries with "Don’t know/Not Sure", "Refused" and missing values, our study may be subject to non-response bias. Additionally, our Poisson regression failed the overdispersion assumption, which indicates that the resulting model will be less accurate, leading to underestimated (lower than the true value) standard error of parameter estimation and, therefore, invalid conclusions.


\newpage
# Appendix

```{r echo = FALSE,results='asis',warning = FALSE,include=TRUE}
# final model result
stargazer(model3,title="log expected counts", align=TRUE,header = FALSE,ci = TRUE,digits = 2,single.row = TRUE,type = 'latex')
```



```{r echo = FALSE,warning=FALSE, message=FALSE,include=TRUE}
# vif table
knitr::kable(cbind(vif[,3]),digits = 2,caption = '$GVIF^{(1/(2*Df))}$')%>% kable_styling(position="center",full_width = FALSE)
```
